---
title: "Fluorcam data analysis - Fv/Fm multiple assays"
author: ""
date: "`r format(Sys.time(), '%m/%Y')`"
output:
  html_document:
    theme: cerulean
    highlight: tango
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
  pdf_document:
    toc: yes
    toc_float: true
    number_sections: yes
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script perform analysis of Fv/Fm data from multiple assays, extracted from fluorcam output files

# Variables

Before execution it is necessary to define variables for the analysis

```{r message=FALSE, echo=FALSE, warning= FALSE}
#Define Variables
##Reference line : the line againt which the statistical tests will be performed
RefLine <- "WT1"  #a string between ""

## Order in wich the lines will appear on the plot
target_order <- c("WT1", "h1", "npq1", "npq4", "h1npq1", "h1npq4") # name of your lines

# Working directory
WORKDIR <- "~/MyData/fluorcam"

```

**Reference line : the line againt which the statistical tests will be performed**<br>
`r {"RefLine: "}`
`r RefLine`<br>

**Order in wich the lines will appear on the plot**<br> 
`r {"target_order: "}`
`r target_order`<br>

**Working directory : The directory where the files are**<br> 
`r {"WORKDIR: "}`
`r WORKDIR`<br>

```{r message=FALSE, echo=FALSE, warning= FALSE}
#==============================================================================
#Installation des packages
if (!require(tidyr)) { install.packages("tidyr") }
if (!require(dplyr)) { install.packages("dplyr") }
if (!require(tidyverse)) { install.packages("tidyverse") }
if (!require(devtools)) { install.packages("devtools") }
if (!require(rstatix)) { install.packages("rstatix", repos = "https://cloud.r-project.org") }
if (!require(ggbeeswarm)) { install.packages("ggbeeswarm") }
if (!require(RColorBrewer)) { install.packages("RColorBrewer") }
if (!require(rcompanion)) { install.packages("rcompanion") }
if (!require(svglite)) { install.packages("svglite") }

################################################################################################################################

# Fonction utilisée dans le script

#=======================================================================================================================================
# Check data normality. Get out of th loop if at least one of the group of data does not follow a normal law.
# return TRUE if data follow a normal law
#=======================================================================================================================================
check_normality <- function(shapiro_df) {
  # The normality is assumed to be true
  flag_normal <- TRUE
  
  for (i in 1 : nrow(shapiro_df)) {
    if(shapiro_df[i, 4] > 0.05) {
      # print(paste0("les données ",shapiro_df$grouping_factor[i],"-", 
      # shapiro_df$plant_line[i], " suivent une loi normale"), quote = FALSE)
      
    } else {
      # print(paste0("les données ",shapiro_df$grouping_factor[i],"-", 
      #          shapiro_df$plant_line[i], " ne suivent pas une loi normale"), quote = FALSE)
      
      # If data don't follow a normal law, don't go further
      flag_normal <- FALSE
      break
    }
  }
  return(flag_normal)
}

#=======================================================================================================================================
# Take the results of Anova test
# Return TRUE if at least one mean is significantly different 
#=======================================================================================================================================
check_anova <- function(anova_results) {
  flag_anova <- FALSE
  
  for (i in 1 : nrow(anova_results)) {
    if (anova_results$p[i] < 0.05) {
      #print (paste0("Anova test compares the mean, pvalue for the group ", anova_results$time[i] , " is < 0.05 this means that at least one of the medians is significantly different from the others, a post hoc Tukey test is performed"))
      flag_anova <- TRUE
    } else {
      #print (paste0("Anova test compares the mean, pvalue for the group", anova_results$time[i] , " is > 0.05 this means that there is no significant difference between means"))
    }
  }    
  
  return(flag_anova)
}

#=======================================================================================================================================
# Perform Kruskal-Wallis test
# Return TRUE if one of the median is significantly different
#=======================================================================================================================================
check_kruskal <- function(kruskal_pval) {
  flag_kruskal <- FALSE
  
  for (i in 1 : nrow(kruskal_pval)) {
    if (kruskal_pval$p[i] < 0.05) {
      #print (paste0("Kruskall Wallis test compares medians, pvalue for the group ", kruskal_pval$time[i] , " is < 0.05 this means that at least one of the medians is significantly different from the others, a post hoc Dunn test is performed"))
      flag_kruskal <- TRUE
    } else {
      #print (paste0("Kruskall Wallis test compares medians, pvalue for the group ", kruskal_pval$time[i] , " is > 0.05 this means that there is no significant difference between medians"))
    }
  }    
  
  return(flag_kruskal)
}

#=======================================================================================================================================
# Perform Dunn test
# Return the pvalue in a dataframe
#=======================================================================================================================================
test_dunn <- function() {
  pval <- as.data.frame(df_data_long%>%
                          mutate(line = fct_relevel(line, target_order)) %>% 
                          group_by(time) %>% dunn_test(value ~ line, p.adjust.method = "BH"))
  #print(df_data_long %>% group_by(time) %>% dunn_test(value ~ line, p.adjust.method = "BH"))
  return(pval)
}

#==============================================================================
```

# Data

The output files from the fluorcam are porcessed by the corresponding script.
The data to load into R-studio looks like :

T0	H1	0.764181557359925	0.770512875444535	0.76009601614817	0.766746075196425  
T0	H1npq1	0.793915032414362	0.809684308883401	0.801904304047648	0.813511707977184  
T0	H1npq4	0.767414953153493	0.770888770622672	0.755604534005038	0.768396025117357  
T0	WT1	0.846261579983952	0.851513191251289	0.849351569697399	0.859264705882353  
T0	npq1	0.847455697103409	0.840633935136018	0.850742288203396	0.856813871579518  

All files to anlyse together must be in the same repertory and named following the pattern :
results_fvfm....TXT  

After loading, the data frame looks like 
```{r message=FALSE, echo=FALSE, warning= FALSE}
#Define Working directory
setwd(WORKDIR)

# # Identify and load the data
# Files are all in the same repertory and are named according to 
# the following pattern : 
# results_fvfm....

temp <- list.files(pattern="^results_fvfm")
myfiles <- lapply(temp, read.table, header = FALSE, sep = "\t")
names(myfiles) <- temp

#Prepare datas
# Combine dataframe without duplicating common columns (columns 1 et 2)
df_data <- do.call(cbind, c(myfiles[1], lapply(myfiles[2:length(myfiles)], '[', -c(1,2))))
colnames(df_data) <- c("time", "line", (paste("rep", c(1:(ncol(df_data)-2)), sep = '')))

knitr::kable(head(df_data),
             caption = "First rows of the data")

#Convert to long format
library(tidyr)
library(tidyverse)
#https://tidyr.tidyverse.org/reference/pivot_longer.html
df_data_long <- df_data %>% pivot_longer(cols = starts_with("rep"), names_to = "replicat")

```

# Statistical analysis

## Checking normality status of the data
```{r message=FALSE, echo=FALSE, warning= FALSE}
library(ggbeeswarm) # for funcnction geom_quasirandom
library(RColorBrewer) # to define colors
library(rstatix)  # for statistical test
library(rcompanion) # to compute confidence interval for non parametric data
library(forcats)
library(svglite)


# Convert line's name into factor to force the order in the plot 
df_data_long$line <- as.factor(df_data_long$line)


# Determining data normality status
shapiro_df <- df_data_long %>%
  dplyr::group_by(time, line) %>%
  summarise(statistic = shapiro.test(value)$statistic, 
            p.value = shapiro.test(value)$p.value)

flag_normal <- check_normality(shapiro_df)
```

<br>  
A shapiro test is performed to check the normality status of the datas

`r {knitr::kable(shapiro_df, caption = "Shapiro test results.", digits = 30)}`

`r if(flag_normal == TRUE) {"**Datas follow a normal law**"} else {"**Datas don't follow a normal law**"}`  

## Pairwise comparaison
```{r message=FALSE, echo=FALSE, warning= FALSE}
# Data treatement according to normality status
if(flag_normal == TRUE) {
  #print("Datas follow a normal law")
  
  # Summary
  if (!require(Rmisc)) {install.packages("Rmisc")}
  library(plyr) # rmisc dependance
  library(Rmisc) # for the command summarySE
  my_summary <- summarySE(df_data_long, measurevar="value", groupvars=c("time", "line"))
  
  # The 2 packages must be detach because they can prevent the goup_by function to work properly
  detach(package:Rmisc)
  detach(package:plyr)
  
  
  # Stats
  anova_results <- df_data_long %>% group_by(time) %>% anova_test(value ~ line)
  flag_anova <- check_anova(anova_results)
  if (flag_anova == TRUE) {
    tukey_results <- as.data.frame(df_data_long %>% mutate(line = fct_relevel(line, target_order)) %>%
                                     group_by(time) %>%  tukey_hsd(value ~ line))
  }
  
  # Save the files
  write.table(my_summary, file = "Summary.txt", 
              quote = FALSE, row.names = FALSE, sep = '\t')
  write.table(anova_results, file = "Anova.txt", 
              quote = FALSE, row.names = FALSE, sep = '\t')
  if (flag_anova == TRUE) {
    write.table(tukey_results[, c(1,3,4, 9, 10)], file = "Tukey.txt", 
                quote = FALSE, row.names = FALSE, sep = '\t')
  } 
  
  # Prepare data to add stars on error bar
  if(flag_anova == TRUE) {
    df <- dplyr::filter(tukey_results, grepl(RefLine, group1))
    
    for (i in levels(as.factor(df$time))) {
      x <- c(i, "line", RefLine, RefLine, rep(x="", 6))
      df <- rbind(x, df)
    }
    
    df <- df[ with(df, order(time, group2)),]
    
    df2 <- cbind(my_summary, df$p.adj.signif)
    names(df2)[names(df2) == "df$p.adj.signif"] <- "Labels"
    
    # Plot
    p<-df2 %>% mutate(line = fct_relevel(line, target_order)) %>%  
      ggplot(aes(x=time, y=value, group = line))+
      geom_col(aes(x=time, y=value), color="dark grey", fill="white")+
      geom_quasirandom(data = df_data_long, aes(x = time, y = value, color = line), 
                       color="green", width = 0.3, alpha = 0.6)+
      geom_errorbar(aes(x = as.factor(time), ymin=value-ci, ymax=value+ci), 
                    width=.2, position=position_dodge(0.9), color = "black") +
      geom_text(aes(x = as.factor(time), y= value+ci + 0.018 , 
                    label = Labels),
                size = 3, position = position_dodge(0.9), inherit.aes = TRUE)+
      facet_wrap(~line, nrow=1)+
      theme_classic()
    
    print(p)
    
    ggsave("fvfm_plot.svg", width=8, height=3)
    
  }else {
    print("Datas are not significantly different, The Tukey test was not performed.")
  }
  
  
} else {
  #print("Datas don't follow a normal law")
  
  # Summary
  conf_int <- groupwiseMedian(data = df_data_long,
                              var = "value",
                              group = c("time", "line"),
                              conf       = 0.95,
                              R          = 5000,
                              percentile = TRUE,
                              bca        = FALSE,
                              digits     = 3)
  
  
  
  # Stats
  kruskal_pval <- (df_data_long %>% group_by(time)%>%kruskal_test(value ~ line)) %>% select(time, p)
  
  flag_kruskal <- check_kruskal(kruskal_pval)
  if (flag_kruskal == TRUE) { pval_dunn <- test_dunn() }
  
  # Save the files
  write.table(conf_int, file = "Summary.txt", 
              quote = FALSE, row.names = FALSE, sep = '\t')
  write.table(kruskal_pval, file = "Kruskal.txt", 
              quote = FALSE, row.names = FALSE, sep = '\t')
  
  if (flag_kruskal == TRUE) {
    write.table(pval_dunn[, c(1, 3, 4, 8, 9, 10)], file = "Dunn.txt", 
                quote = FALSE, row.names = FALSE, sep = '\t')
  } 
  
  # Prepare data to add stars on error bar
  if(flag_kruskal == TRUE) {
    df <- dplyr::filter(pval_dunn, grepl(RefLine, group1))
    
    for (i in levels(as.factor(df$time))) {
      x <- c(i, "line", RefLine, RefLine, rep(x="", 6))
      df <- rbind(x, df)
    }
    
    df <- df[ with(df, order(time, group2)),]
    
    df2 <- cbind(conf_int, df$p.adj.signif)
    names(df2)[names(df2) == "df$p.adj.signif"] <- "Labels"
    
    # Plot
    p<-df2 %>% mutate(line = fct_relevel(line, target_order)) %>%  
      ggplot(aes(x=time, y=Median, group = line))+
      geom_col(aes(x=time, y=Median), color="dark grey", fill="white")+
      geom_quasirandom(data = df_data_long, aes(x = time, y = value, color = line), 
                       color="green", width = 0.3, alpha = 0.6)+
      geom_errorbar(aes(x = as.factor(time), ymin=Percentile.lower, ymax=Percentile.upper), 
                    width=.2, position=position_dodge(0.9), color = "black") +
      geom_text(aes(x = as.factor(time), y= Percentile.upper + 0.018 , 
                    label = Labels),
                size = 3, position = position_dodge(0.9), inherit.aes = TRUE)+
      facet_wrap(~line, nrow=1)+
      theme_classic()
    
    print(p)
    
    ggsave("fvfm_plot.svg", width=8, height=3)
    
  }else {
    print("Datas are not significantly different, The Dunn test was not performed.")
  }
}
```
<br>  
`r if(flag_normal == TRUE) {"Anova test is performed"} else {"Kurskal-Wallis test is performed"}`  
`r if(flag_normal == TRUE) {knitr::kable(anova_results, caption = "Anova test results")}`
`r if (exists("flag_anova")) {if (flag_anova == TRUE) {knitr::kable(tukey_results, caption = "Tukey post hoc test results")} else { "Differences between the datas are not significant"}}`

`r if(flag_normal == FALSE) {knitr::kable(kruskal_pval, caption = "Kruskal-Wallis test results")}`
`r if (exists("flag_kruskal")) {if (flag_kruskal == TRUE) {"Kruskall Wallis test compares medians, pvalue for at least one of the group is < 0.05 this means that at least one of the medians is significantly different from the others, a post hoc Dunn test is performed"}}`
`r if (exists("flag_kruskal")) {if (flag_kruskal == TRUE) {knitr::kable(pval_dunn, caption = "Dunn post hoc test results")} else { "Differences between the datas are not significant"}}`  

`r {"symbols = **** pval < 1e-04 , *** pval < 0.001, ** pval < 0.01, * pval < 0.05, ns pval > 0.05"}`  
<br>  

# Bibliographie  

Package	| Version |	Citation
--------|---------|---------
base	| 4.2.2	| R Core Team (2022)
devtools |	2.4.5 |	Wickham, Hester, et al. (2022)
ggbeeswarm	| 0.6.0 |	Clarke and Sherrill-Mix (2017)
grateful |	0.1.11 |	Rodríguez-Sánchez, Jackson, and Hutchins (2022)
knitr |	1.40 |	Xie (2014); Xie (2015); Xie (2022)
plyr |	1.8.7 |	Wickham (2011)
RColorBrewer |	1.1.3 |	Neuwirth (2022)
rcompanion |	2.4.18 |	Mangiafico (2022)
rmarkdown	| 2.18 |	Xie, Allaire, and Grolemund (2018); Xie, Dervieux, and Riederer (2020); Allaire et al. (2022)
Rmisc	| 1.5.1 |	Hope (2022)
rstatix	| 0.7.0 |	Kassambara (2021)
svglite	| 2.1.0 |	Wickham, Henry, et al. (2022)
tidyverse	| 1.3.2 |	Wickham et al. (2019)  

# Information session R
```{r echo= FALSE}
#Environnement
devtools::session_info()
```
