---
title: "Fluorcam data analysis - NPQ"
author: ""
date: "`r format(Sys.time(), '%m/%Y')`"
output:
  html_document:
    theme: cerulean
    highlight: tango
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
  pdf_document:
    toc: yes
    toc_float: true
    number_sections: yes
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script perform analysis of NPQ data, extracted from fluorcam output files

# Variables

Before execution it is necessary to define variables for the analysis

```{r message=FALSE, echo=FALSE, warning= FALSE}
#Define Variables
# Working directory
WORKDIR <- "~/MyData/fluorcam"

# Order in wich the lines will appear on the plot
target_order <- c("WT1", "h1", "npq1", "npq4", "h1npq1", "h1npq4")   #your lines names

# File to convert code in the fluorcam output file into length of mesure
CODE <- "your_time_point_file.txt"
```

**Order in wich the lines will appear on the plot**<br> 
`r {"target_order: "}`
`r target_order`<br>

**Working directory : The directory where the files are**<br> 
`r {"WORKDIR: "}`
`r WORKDIR`<br>


```{r message=FALSE, echo=FALSE, warning= FALSE}
#==============================================================================
#Installation des packages
if (!require(tidyr)) { install.packages("tidyr") }
if (!require(dplyr)) { install.packages("dplyr") }
if (!require(rcompanion)) { install.packages("rcompanion") }
if (!require(ggplot2)) { install.packages("ggplot2") }
if (!require(svglite)) { install.packages("svglite") }

library(dplyr)
library(rcompanion)
library(ggplot2)
library(svglite)

################################################################################################################################

# Fonction utilisée dans le script

#=====================================================================================================================================
#Check data normality. Get out of th loop if at least one of the group of data does not follow a normal law.
# return TRUE if data follow a normal law
#=====================================================================================================================================
check_normality <- function(shapiro_df) {
  # The normality is assumed to be true
  flag_normal <- TRUE
  
  for (i in 1 : nrow(shapiro_df)) {
    if(shapiro_df[i, 4] > 0.05) {
      # print(paste0("les données ",shapiro_df$grouping_factor[i],"-", 
      # shapiro_df$plant_line[i], " suivent une loi normale"), quote = FALSE)
      
    } else {
      # print(paste0("les données ",shapiro_df$grouping_factor[i],"-", 
      #          shapiro_df$plant_line[i], " ne suivent pas une loi normale"), quote = FALSE)
      
      # En fait les données ne sont pas normales, pas besoin d'aller plus loin
      flag_normal <- FALSE
      break
    }
  }
  return(flag_normal)
}

#=====================================================================================================================================
# Perform analysis
# Return the plot
#=====================================================================================================================================
f_analyseNPQ <- function(df){
  #Determining data normality status
  
  shapiro_df <- df %>%
    dplyr::group_by(line, secondes) %>%
    summarise(statistic = shapiro.test(value)$statistic,
              p.value = shapiro.test(value)$p.value)
  
  flag_normal <- check_normality(shapiro_df)
  
  # Data treatement according to normality status
  if(flag_normal == TRUE) {
    #print("Datas follow a normal law")
    
    # Summary
    if (!require(Rmisc)) {install.packages("Rmisc")}
    library(plyr) # dépendence de rmisc
    library(Rmisc) # pour la commande summarySE
    my_summary <- summarySE(df, measurevar="value", groupvars=c("line", "secondes"))
    
    detach(package:Rmisc)
    detach(package:plyr)
    
    # Plot
    
    p<-my_summary %>%
      ggplot(aes(x=secondes, y=value, group = line))+
      geom_ribbon(aes(ymin=value-ci, ymax=value+ci, fill=line),alpha=0.3, linetype=0)+
      scale_fill_manual(values=my_colours)+
      geom_point(aes(x=secondes, y=value, color=line))+
      geom_line(aes(x=secondes, y=value, color=line), size=1)+
      scale_colour_manual(values=my_colours) +
      facet_wrap(~fct_relevel(line, target_order))+
      theme_classic()+
      theme(legend.position= "none")+
      ggtitle(levels(as.factor(df$time)))+
      theme(plot.title=element_text(hjust = 1))
      
    
    
    
  } else {
    #print("Datas don't follow a normal law")
    
    # Summary
    conf_int <- groupwiseMedian(data = df,
                                var = "value",
                                group = c("line", "secondes"),
                                conf       = 0.95,
                                R          = 5000,
                                percentile = TRUE,
                                bca        = FALSE,
                                digits     = 3)
    
    
    # Plot
    p<-conf_int %>%   
      ggplot(aes(x=secondes, y=Median, group = line))+
      geom_ribbon(aes(ymin=Percentile.lower, ymax=Percentile.upper, fill=line),alpha=0.3, linetype=0)+
      scale_fill_manual(values=my_colours)+
      geom_point(aes(x=secondes, y=Median, color=line))+
      geom_line(aes(x=secondes, y=Median, color=line), size=1)+
      scale_colour_manual(values=my_colours) +
      facet_wrap(~fct_relevel(line, target_order))+
      theme_classic()+
      theme(legend.position= "none")+
      ggtitle(levels(as.factor(df$time)))+
      theme(plot.title=element_text(hjust = 1))
    
    
  }
  return_list <- list(flag_normal, plot = p)
  return(return_list)
}

```

# Data

The output files from the fluorcam are processed by the corresponding script.
The data to load into R-studio looks like :

T0	WT1	L1	0.541398766566068	0.538327175008398	0.530264520380307	0.513742842269651  
T0	WT1	L2	1.2385187067268	1.21172606973824	1.23726947426369	1.21259273349819  
T0	WT1	L3	1.55499166243747	1.5449762892709	1.57888474656937	1.54153738583228  
T0	WT1	L4	1.69488414774031	1.69991746256338	1.7440918298447	1.70477164914892  
T0	WT1	L5	1.73493461642932	1.74918957858086	1.82038603339839	1.78241316620419  
T0	WT1	L6	1.7012877510348	1.71003629477671	1.78225150827949	1.74433748584372  
T0	WT1	L7	1.69333944743781	1.69409365440427	1.73854447439353	1.69812108559499   


```{r message=FALSE, echo=FALSE, warning= FALSE}
#Define Working directory
setwd(WORKDIR)

# Identify and load the data
# Files are all in the same repertory and are named according to 
# the following pattern : 
# results_npq....

temp <- list.files(pattern="^results_npq")
myfiles <- lapply(temp, read.table, header = FALSE, sep = "\t")
names(myfiles) <- temp

#Prepare datas
# Combine dataframe without duplicating common columns (columns 1 et 2)
df_data <- do.call(cbind, c(myfiles[1], lapply(myfiles[2:length(myfiles)], '[', -c(1:3))))
colnames(df_data) <- c("time", "line", "secondes", (paste("rep", c(1:(ncol(df_data)-3)), sep = '')))

#Convert to long format
library(tidyr)
library(tidyverse)
#https://tidyr.tidyverse.org/reference/pivot_longer.html
df_data_long <- df_data %>% pivot_longer(cols = starts_with("rep"), names_to = "replicat")

# Load Code for length of mesurement
code <- read.table(CODE, header = TRUE, sep = '\t')

knitr::kable(code,
             caption = "Conversion code") 

# Replace value in column "secondes" according to the code
df_data_long$secondes <- code$secondes[match(df_data_long$secondes, code$code)]



```
<br>
After loading, the data frame looks like
```{r message=FALSE, echo=FALSE, warning= FALSE}
knitr::kable(head(df_data_long),
             caption = "First rows of the data")

# Convert line's name into factor to force the order in the plot  
df_data_long$line <- as.factor(df_data_long$line)
 

# Split the data frame into a list of data frame, one per line
df_list <- split(df_data_long, f = df_data_long$time)

# Define color
my_colours = rep("chartreuse3", each = length(levels(as.factor(df_data$line))))
```

# Results  

A shapiro test is performed to check the normality status of the datas
```{r message=FALSE, echo=FALSE, warning= FALSE}
# Perform analysis
result <- lapply(df_list, f_analyseNPQ)

ggsave("npq_plot.svg", width=8, height=3)
```

`r if(result[[1]][[1]] == TRUE) {"**TO Datas follow a normal law**"} else {"**T0 Datas don't follow a normal law**"}`  

```{r message=FALSE, echo=FALSE, warning= FALSE}
result[[1]]$plot
```

`r if(result[[2]][[1]] == TRUE) {"**T24 Datas follow a normal law**"} else {"**T24 Datas don't follow a normal law**"}`  

```{r message=FALSE, echo=FALSE, warning= FALSE}
result[[2]]$plot
```

`r if(result[[3]][[1]] == TRUE) {"**T48 Datas follow a normal law**"} else {"**T48 Datas don't follow a normal law**"}`  

```{r message=FALSE, echo=FALSE, warning= FALSE}
result[[3]]$plot
```

# Bibliography   

Package	| Version |	Citation
--------|---------|---------
base	| 4.2.2	| R Core Team (2022)
devtools |	2.4.5 |	Wickham, Hester, et al. (2022)
grateful |	0.1.11 |	Rodríguez-Sánchez, Jackson, and Hutchins (2022)
knitr |	1.40 |	Xie (2014); Xie (2015); Xie (2022)
plyr |	1.8.7 |	Wickham (2011)
rmarkdown	| 2.18 |	Xie, Allaire, and Grolemund (2018); Xie, Dervieux, and Riederer (2020); Allaire et al. (2022)
Rmisc	| 1.5.1 |	Hope (2022)
svglite	| 2.1.0 |	Wickham, Henry, et al. (2022)
 

# Information session R
```{r echo= FALSE}
#Environnement
devtools::session_info()
```
  



